Fish do not need Visas: the transboundary nature of the world’s exploited marine fish species

### 2019-12-31 Manuscript_Draft.Rmd
Finished second draft and started abstract

---
title: "Fish do not need Visas: the transboundary nature of the world’s exploited marine fish species"
author: "Authors: Juliano Palacios-Abrantes^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada, ^2^Yale University, New Haven, Connecticut, USA, ^3^Stockholm Resilience Centre, Stockholm University, Stockholm, Sweden"
subtitle: "DRAFT DO NOT SHARE"
# subtitle: "Fish do not need Visas (16/40 character maximum)"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
# bibliography: ./References/FishForVisa_Ref.bib
# csl: ./References/science.csl
---

@@

@@

Distributions of marine species around the world are not bounded by human-made boundaries; rather shaped by contemporary biotic and abiotic factors, as well as their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. However, management of living marine resources is predicated on the definition of “stocks” that are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, established political boundaries across the distribution of many species, creating shared species between nations. Theories and empirical experience demonstrate fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@REF]. For example, evironmental fluctuation leading to changes in the stock distribution [@Munro:2002uf] and fishing quotas set by different nations for a shared fishery resource have led to conflict [@Spijkers:2017ij]. Moreover, climate change, which is shifting species distributions, is likely to exacerbate this type of conflict [@Pinsky:2018cb] and presents further challenges for fisheries management [@Miller:2013iv].

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that distribute across two or more EEZs (hereafter called shared species) in the world is still unclear. Caddy in 1997 estimated that there could be up to 1,500 transboundary fish species in the world. However, such estimation lacked a proper assessment due to limited information on the distribution of species at the time [@Caddy:1997ue]. More recently, Teh et al. (2010) estimated that 206 species can be considered shared, accounting for a total catch of 34.2 million tones and a global landed value of USD 30.7 trillion. While this represents a good starting point, the analysis was based on a literature review of shared species and did not include a mechanistic way of determining the transboundary nature of the species [@Teh:2015gd]. Getting accurate estimates of the number of shared species and where they are located can help develop fisheries management and conservation plans that account for the challenges from transboundary resources management [@Millis].

<!-- ## Methods summary (250/250 words) -->

This paper aims to identify the world current transboundary species and determine their contribution to the catch and revenue of the world marine fisheries. The Food and Agricultural Organization (FAO) recognizes three types of shared stocks: (i) transboundary stocks, shared by 2 or more neighboring coastal nations; (ii) straddling stocks, occurring in two or more adjacent national jurisdictions and the high seas; and (iii) highly migratory stocks, found in the EEZs of coastal nations that are not necessarily adjoining, and the high seas. We focus on estimating the number of species that meet the first category. We overlay the known distribution of **968** commercially valuable marine fish species and the Exclusive Economic Zones of 153 coastal countries to determine the total number of transboundary species and their contribution to global catch and landed value. We carry out the analysis at the species level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species [@Teh:2015gd].

We used four-criteria that combined multiple species distribution data-sources and the world's EEZs to determine if a species can be considered a “shared species”. The first criteria consisted in only considering shared species between neighbouring EEZs, rather than the extended distribution. For the second criteria we considered a species to be present in a grid cell if all data sources predicted positive occurrence. Current species distributions were predicted from occurrence records obtained from publicly available observational databases of the world’s marine species and two structurally distinct species distribution models [@Asch:2018ca; Reygondeau:2019uh; @Zeller2016] (see Methods). The third criteria determined whether the species were exploited at the grid cell using spatially explicit catch data [@Zeller2016]. A species was kept if both catch data and predicted species occurrences were positive at the same spatial cells. Finally, the last criteria determined a spatial treshold to confirm that a species occurs in both neighboring EEZs. Once we have identified the shared stocks, we estimated the contribution of the species to fishing revenues and landings using the SAU database [@Zeller2016] and categorized them according to catch trends [@Kleisner:2013gh].

## Results and discussion (~500/1000 words)

Overall, we estimated **633** transboundary species in the world, more than double the number previously estimated [@Teh:2015gd], accounting for **65.4**% of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared stocks within EEZs landed an average of **69.2** million tonnes, representing **64**% of the EEZ-global catch, and generating USD **79,745** million in fishing revenue. This is equivalent to **49.8**% of the global revenue of the same period. These values are higher than previously estimated reinforcing the importance of these stocks at the global level [@Teh:2015gd].

```{r Call_Map, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="A) Number of transboundary species per EEZ and they contribution to countries fishing revenue. The number of species is represented in the EEZ while the revenue is represented in the land polygon. B) Regional revenue (left) and catch (right) contribution of tranbsoundary species.Points = mean +- sd."}

@@

![**A) Number of transboundary species per EEZ and they contribution to countries fishing revenue. The number of species is represented in the EEZ while the revenue is represented in the land polygon. B) Regional revenue (left) and catch (right) contribution of tranbsoundary species. Points = mean +- sd.**](./Figures/Fig1.png)

Transboundary species are particulary important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions is driven by China (USD 6,279 million), in E. Asia and the US (USD 9,605 million) in N. America who contribue 55% and 83% of the region's revenue, respectiveley. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species. These 5 nations are responsable for 30% of the global revenue from transboundary species. Peru and Russia lead the production of transboundary species harvesting over 5,000 thousand tones (**Suplemental figure Trans_Catch_MAP**). Peru's catches come mainly from Peruvian anchoveta (*Engraulis ringens*), shared with Chile, accounting for 79% of the transboundary species production. Similarly, Russia shares the (currently) largest fishery in the world, Alaskan polllock (*Theragra chalcogramma*) with both Japan and the US [@FAO:2018KK]. Despite this, catches are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@FAO:2018KK] getting around 1/3 of its capture from transboundary species, followde by Japan (4.2 million tonnes) and South Korea (2.3 million tonnes).

The exploitation category of transboundary species significantly differs from each other (one-way ANOVA *F*(5,446), *f* = 324.5, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05 for all) with most species fished under 50% of their historical maximum peak (Fig.2). This is true for all (but Oceania) continents with no significan difference between them (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05), suggesting that the current global exploitation rate of transboundary species could be over the maximum sustainable yield [@Kleisner:2011wn; @Pauly:2013gs]. Countries that are today harvesting at this rate (Category C) should be paying extra care in the management of these species (Fig. 2).

@@

While these catch trends do not directly reflect stock status - e.g. other factors like markets, fishing policies or extreem events also influence catch volume [@Pauly:2013gs; @Branch:2008kk] - they could ressemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. Empirical analysis suggest that in most cases, management of transboundary species will have better outcomes in terms of fish yield and sustainabillity whith cooperation between nations [@Bailey:2010ga]. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*) [@Diekert:2010gp], and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Collaboration can also reduce conflict between nations over shared ressources. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts like Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations to compensate losses [@Miller2003]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for Atlantic mackerel (*Scomber scombrus*) [@Spijkers:2017ij], a species that is transboundary and contributes **698.3** thousand tones to landings and USD **968** million in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change could exacerbate conflicts over transboundary species as climate change shifts the distribution of marine species around the globe [@Pinsky:2018cb].

First steps for shared management start with information sharing, then this and that. Proper management treaties not only have but can also be stronger in the fase of a canging climate.

Species can have a cosmopolitan distribution, that is, a broader distribution, or be limited to a specific region, known as endemic [@Nelson:2016kk]. Not surprisingly, highly migratory pelagic-oceanic species are the only group that significantly differs in terms of sharing countries (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 nations shared per species (Fig. 3).

@@

Due to their large migration patterns and presence in areas beyond national jurisdiction, pelagic-oceanic species such as skipjack tuna (*Katsuwonus pelamis*), Common thresher shark (*Alopias vulpinus*), and tunas (*Thunnus sp.*) are catalogued as stradling species and globally managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species, most of them fished in areas beyond national jurisdiction [@CullisSuzuki:2010fi]. We estimate that 81.8% of these species (*n* = 81) are actually shared within neighbouring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighbouring RFMOs (**Suplemental figure RFMO_MAP**). The catch from pelagic-oceanic species comes mainly from five South Pacific Islands nations, from which three (Kiribalti, Solomon and Papua New Guinea) are members of the Pacific Nauru Agreement (PNA) (Fig. 3B). The PNA manages 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year [@https://www.ffa.int/nauru_agreement].

The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these might be cosmopolitan but not necessarily highly migratory (e.g. *Solea solea*, *Panulirus argus*) (Fig. 3). However, less distributed species must not be ignored. The world of fisheries is highly connected trough larvae dispersal [@Ramesh:2019va] and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. Moreover, many coastal communities, specially in developing regions, and highly dependent on fishing local fish (e.g. reef fish) for food security and improved livelihoods [@CisnerosMontemayor:2016gq; Wabnitz:2018gf; @https://www.sciencedirect.com/science/article/pii/S0308597X17308357]. For example, from the top five countries fishing transboundary reef - associated species, Indonesia, Malaysia and the Philippines are among the higest fish consumption per capita in the world with 20, 60 and 40 kg per year, respectiveley [@https://www.sciencedirect.com/science/article/pii/S0308597X12001315].

The global contribution of transboundary species to fishing revenue and capture is compelling, with transboundary species playing an important role in many fishing nations. Transboundary species have played and continue to play an important role in fisheries conflicts, with climate change likely to exacerbate this trend given expected shifts in species distribution and the emergence of 'new' shared species. Identifying existing transboundary species is the first step towards joint management frameworks that are precautionary, strive for sustainability and can be flexible to accommodate future changes.

@@

The previous models combine observational data with a series of biotic and abiotic information to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g. ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - i.e., catches  based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in the FAO data available for each country. Catches are also spatially allocated [@Zeller2016]. Thus, we used the Sea Around Us catch reconstruction database from 2005 to 2014 (last ten yeares of data) as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs.

@@

We developed a four-criteria method for determining whether or not a species was transboundary. Only species that meet all criteria were considered as transboundary species and the analysis was done only within the EEZs of coastal states. In some cases, criteria also work as indexes to measure the uncertainty in the analysis as they provided continuos index (e.g. from 0 to 1) rather than a discrete, yes or no, index . All of the analysis was done in the statistical software *R version 3.5.2 (2018-12-20)* with the pacages *tidiverse*[@PackagetidyverseE:2017vq], *data.table*[@Packagedatatable:2019uh], *wesanderson*[@], *janitor*[@], *spdep*[@], *sf*[@PackagesfSimpleF:2018vp], *sp*[@], *rgdal*[@PackagergadalBind:2018vy], *tools*[@], *parallel*[@], *rfishbase*[@], *zoo*[@]. All code is available at https://github.com/jepa/FishForVisa.

@@

We use the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the 0.5 x 0.5 grid cells. We computed a *Species Index* by dividing the number of datasets that confirm presence of a species in a grid cell over the total amount of datasets (*n* = 3). The index is unit-less and represents a scale from 0.3 to 1 where 0.3 means only one dataset reports presence of the species in a grid cell, 0.6 represents two of three sources, and 1 means agreement between all datasets. For the analysis, we selected only those cases where *Species Index* = 1 to get a more conservative estimate of transboundary species (**S1_Map_per_dataset**).

@@

We estimated the economic contribution of transboundary species for each country using global ex-vessel price [@Tai2017]. The data includes ex-vessel price from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The contribution of transboundary species was done by aggregating the catch of each species that happened within the EEZ of the countries sharing that species. For this study, we did not include catch from areas beyond national jurisdiction. Finally, we use the catch data to determine the exploitation category of each species within a EEZ. Although this method has previously used to estimate stock status [@Grainger:1996tk; @Kleisner:2011wn], the categories presented here are intended to be seen as catch trends, and not status of each species. We only assessed species within each EEZ that had at least ten years of data between the first and last reported landings with at least five consecutive data years. Three categories were drawn from the method depending on the year's catch, the maximum catch over time and the minimum catch after the catch peak (**Table XX**). Finally, we reported the predominant category over the last ten years of catch data (2005-2014).

```{r Status_Table, eval = F, echo = F}

@@

Stock_Status <- c("A - Growing Fishery","B - Stable Fishery"," C - Fishery under 10% of Peak","No Status")

@@

data.frame(Stock_Status,
           Rules) %>% 
  kable("latex", booktabs = T,
        caption = "Rules to determine the status of each transboundary species") %>%
  kable_styling(full_width = F) %>%
  column_spec(1,width = "3cm") %>% #wraps text in column2
  column_spec(2,width = "13cm") #%>% #wraps text in column2

@@

# Suplements

@@

**Suplemental figure Trans_Catch_MAP**

# Recicled text (Not to be included)

>>>>

---
title: "Fish do not need Visas: the transboundary nature of the world’s exploited marine fish species"
author: "Authors: Juliano Palacios-Abrantes^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada, ^2^Yale University, New Haven, Connecticut, USA, ^3^Stockholm Resilience Centre, Stockholm University, Stockholm, Sweden"
subtitle: "DRAFT DO NOT SHARE"
# subtitle: "Fish do not need Visas (16/40 character maximum)"
output:
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
  pdf_document:
    fig_caption: yes
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ./References/FishForVisa_Ref.bib
csl: ./References/science.csl
---

@@

The precense of a species in time and space is linked to a combination of biotic and abiotic factors, but the concept of species distribution is rarely integrated into fisheries management, or policy, .

Identifying what species re transboundary is a first step to the proper management of shared stocks and achieving sustainabillity of marine ressources.

Incorporating the concept of species distribution into fisheries management, or policy, is not allways hapens as shown by the delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s. Shared marine ressources are inheratably complex to manage and thus, require special care. However, forty years after the formal adoption of EEZs, the question of how many shared species exist in the world based on their distribution is still unanswered. Here, we ucombine species' distributions with the nation's EEZs in order to determine the number of transboundary stocks. moreover, we identify their importance in terms of catch and revenue and discuss the

@@

Distributions of marine species around the world are not bounded by human-made boundaries; rather shaped by contemporary biotic and abiotic factors, as well as their evolutionary history [@Hutchinson:1957nt; @Nelson:2016kk]. Thus, a species can be cosmopolitan and have a broader distribution like the spiny dogfish (*Squalus acanthias*), or endemic like the masu salmon (*Oncorhynchus masou*) who's distribution is limited to the Japanese coast [@Nelson:2016kk]. However, management of living marine resources is predicated on the definition of “stocks” that are delineated by human-made spatial boundaries that often do not correspond to biologically-meaningful population units [@Song:2017iua; @FredstonHermann:2018kp]. The delineation of Economic Exclusive Zones (EEZs) under the United Nations Convention on the Law of the Seas (UNCLOS) in the early 80s, for instance, established political boundaries across the distribution of many species, creating shared species between nations. Theories and empirical experience demonstrate fisheries targeting resources that straddle between political boundaries complicate fisheries management and potentially reduce its effectiveness to achieve stated objectives [@Bailey:2010ga]. For example, environmental fluctuation leading to changes in species distribution of a shared fishery have led to conflict because of changes in the proportion of captures between countries [@Munro:2002uf], or pressure in the quota allocation due to new participants wanting to join the fishery [@Spijkers:2017ij]. Moreover, climate change, which is shifting species distributions, is likely to exacerbate these types of conflict [@Pinsky:2018cb] and presents further challenges for fisheries management [@Miller:2013iv].

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that are shared by 2 or more neighboring coastal nations (hereafter called transboundary species [@Munro:2002uf]) in the world is still unclear. Caddy in 1997 estimated that there could be up to 1,500 transboundary fish species in the world. However, such estimation lacked a proper assessment due to limited information on the distribution of species at the time [@Caddy:1997ue]. More recently, Teh and Sumaila (2010) estimated that 206 species can be considered shared, accounting for a total catch of 34.2 million tones and a global landed value of USD 30.7 trillion. Such analysis was based on a literature review of shared species and did not include a mechanistic way of determining the transboundary nature of the species [@Teh:2015gd]. In here, we overlay the known distribution of **968** commercially valuable marine fish species and the EEZ of 153 coastal countries to identify those that are shared between neighboring EEZs and determine their contribution to the world's fisheries catch and revenue.

We developed a four-criteria methodology to determine if a species was transboundary (see Methods). The first criteria consisted in only considering shared species between neighboring EEZs, rather than the specie's extended distribution. For the second criteria we considered a species to be present in a $0.5^o$ x $5^o$ grid cell if all data sources predicted positive occurrence. Current species distributions were predicted from occurrence records obtained from publicly available observational databases of the world’s marine species and two structurally distinct species distribution models [@Asch:2018ca; Reygondeau:2019uh; @Zeller2016]. The third criteria determined whether the species were exploited at the grid cell using spatially explicit catch data [@Zeller2016]. A species was kept in a grid cell if both predicted species occurrences and catch data were positive. Finally, the last criteria determined a spatial threshold to confirm that a species occurs in both neighboring EEZs. Once transboundary species were identified, we estimated their contribution to fishing landings and revenues using the *Sea Around Us* database [@Zeller2016] and categorized them according to catch trends [@Kleisner:2013gh]. We carry out the analysis at the species per EEZ level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species [@Teh:2015gd]. Despite loosing detail-specific (e.g. a species could have multiple stocks within an EEZ),  many fisheries within a country or EEZ are managed at the species level, and recent research has suggested strong connectivity between stocks thousand of kilometers away [@Ramesh:2019va].

## Results and discussion (~1000 words)

We identified **633** transboundary species in the world, more than double the number previously estimated [@Teh:2015gd], accounting for **65.4%** of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared species within EEZs landed an average of **69.2** million tonnes, representing **64%** of the EEZ-global catch, and generating USD **79,745** million in fishing revenue. This is equivalent to **49.8%** of the global revenue of the same period. These values are higher than previously estimated reinforcing the importance of these species at the global level [@Teh:2015gd].

```{r Call_Map, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species and they contribution to fishing catch and revenue.** A) The number of species is represented in the EEZ while revenue from these species fished within the EEZ is represented in the land polygon. B) Regional revenue (left) and catch (right) contribution of tranbsoundary species. Points = mean +- sd."}

@@

![**Fig. 1 Number of transboundary species and they contribution to fishing catch and revenue.** A) The number of species is represented in the EEZ while revenue from these species fished within the EEZ is represented in the land polygon. B) Regional revenue (left) and catch (right) contribution of tranbsoundary species. Points = mean +- sd.](./Figures/Fig1.png)

Transboundary species are particularly important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions is driven by China (USD 6,279 million), in E. Asia and the US (USD 9,605 million) in N. America who contribute 55% and 83% to the region's revenue, respectively. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species. These five nations are responsible for 30% of the global revenue from transboundary species. Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead the production of transboundary species harvesting over 5,000 thousand tones (**Supplemental figure Trans_Catch_MAP**). Peru's catches come mainly from Peruvian anchoveta (*Engraulis ringens*), shared with Chile, accounting for 79% of the transboundary species production. Similarly, Russia shares the Alaskan polllock (*Theragra chalcogramma*) fishery with both Japan and the US. Despite this, catches are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@FAO:2018KK] getting around 1/3 of its capture from transboundary species, followde by Japan (4.2 million tonnes) and South Korea (2.3 million tonnes).

Effective management of shared species is a challenging process as it often requires countries to agree on mutual targets in order to achieve sustainable harvest [@Munro:2002uf]. We found significant differences in the exploitation rate of transboundary species (one-way ANOVA *F*(5,446), *f* = 324.5, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05 for all) with most species fished at levels below 50%, sometimes lower than 10%, of their historical maximum peak (Fig.2). This is the case of coastal nations in all continents, but Oceania, with no significant difference between continents (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05), suggesting that the current global exploitation rate of transboundary species could be over the maximum sustainable yield [@Kleisner:2011wn; @Pauly:2013gs]. While these catch trends do not directly reflect stock status - e.g. other factors like markets, fishing policies or extreme events also influence catch volume [@Pauly:2013gs; @Branch:2008kk] - they could resemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. Countries that are today harvesting transboundary species at this rate (Category C) should revise their management policies to make sure these align with sustainable harvest of shared resources [@Munro:2004ug].

@@

Empirical analysis suggest that in most cases, management of transboundary species will have better outcomes in terms of fish yield when nations cooperate [@Bailey:2010ga]. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*) [@Diekert:2010gp; @Eide:2013dha], and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration in shared stocks may threaten stock sustainability [@Clark:1980js; @Nguyen:2018dk], reduce the profitability potential of the fishery [@Merino:2007jz], although cooperation and non-cooperation can be more complex [@Jensen:2015cf]. Collaboration can also reduce conflict between nations over shared resources. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts like Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations to compensate losses [@Munro:2002uf]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for Atlantic mackerel (*Scomber scombrus*) [@Spijkers:2017ij], a species that is transboundary and contributes **698.3** thousand tones to landings and USD **968** million in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change could exacerbate conflicts over transboundary species as climate change shifts the distribution of marine species around the globe [@Pinsky:2018cb].

<!-- First steps for shared management start with information sharing, then this and that. Proper management treaties not only have but can also be stronger in the fase of a canging climate. -->

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], many of which are highly migratory pelagic-oceanic species like skipjack tuna (*Katsuwonus pelamis*), Common dolphinfish (*Coryphaena hippurus*), and tunas (*Thunnus sp.*). This is the only group whoes EEZ range (e.g. the number of EEZs where the species occur as transboundary) significantly differs from the others (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species (Fig. 3).

@@

Due to their large migration patterns and presence in areas beyond national jurisdiction, highly migraroty pelagic fish species are cataloged as straddling and many of them globally managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species, most of them fished in areas beyond national jurisdiction [@CullisSuzuki:2010fi]. We estimate that 81.8% of these species are actually shared within neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighboring RFMOs (**Supplemental figure RFMO_MAP**). The effectiveness of RFMOs responsible for managing high migratory stocks has been previously questioned [@CullisSuzuki:2010fi; Lodge:2007bt]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 13 species (40 stocks) suggests that most tuna and billfish stocks are sustainably harvested ($B/B_{MSY} > 1$ and $F/F_{MSY} < 1$) with only three stocks currently overfished [@Pons:2016ep]. Within national waters, we estimated that the catch of transboundary pelagic-oceanic species comes mainly from five Island Nations in the South Pacific from which three (Kiribalti, Solomon and Papua New Guinea) are members of the Pacific Nauru Agreement (Fig 3B). The parties to the Nauru Agreement sustainably manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

<!-- [@https://www.ffa.int/nauru_agreement]  -->

The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these species have a narrower distribution or are less mobile (e.g. *Solea solea*, *Panulirus argus*) (Fig. 3). However, less distributed species must not be ignored. The world of fisheries is highly connected trough larvae dispersal [@Ramesh:2019va] and so, even centralized populations could be supplying populations thousands of kilometers away [@Ramesh:2019va]. Moreover, many coastal communities, specially in developing regions, and highly dependent on subsistance fishing (e.g. reef fish) for food security and livelihoods [@CisnerosMontemayor:2016gq; Wabnitz:2018gf; @Cabral:2018hc]. For example, from the top five countries fishing transboundary reef - associated species (Fig. 3B), Indonesia, Malaysia and the Philippines are among the highest fish consumption per capita in the world with 20, 60 and 40 kg per year, respectively [@Foale:2013iw]. In these countries, reeef fish contributes substantially to both food security and livelehood of coastal communities [@Cabral:2018hc].

Our study identifies the current species shared by neighbouring coastal nations and highlights the importance of these species in terms of global capture and revenue from wild fisheries. Our results show that captures and revenues from shared species are higher than previously estimated and particularly important to countries in Northern America and Eastern Asia. This result highlights the importance of these fisheries and their potential contribution to regional food security and wellbeing. Moreover, the vast majority of transboundary species are harvest at low rates with no geographic variation. Previous work has found that collaboration will result in better outcomes for shared marine resources. Thus, countries should pay specific attention in the management of the identified transboundary stocks potentially creating collaborative management plans. Identifying existing transboundary species is the first step towards joint management frameworks that are precautionary, strive for sustainability, and can be flexible to accommodate the uncertain future driven by climate change.

@@

The previous models combine observational data with a series of biotic and abiotic information to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g. ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - i.e., catches  based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in the FAO data available for each country. Catches are also spatially allocated [@Zeller2016]. Thus, we used the Sea Around Us catch reconstruction database from 2005 to 2014 (last ten years of data) as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs.

@@

We developed a four-criteria method for determining whether or not a species was transboundary. Only species that meet all criteria were considered as transboundary species and the analysis was done only within the EEZs of coastal states. In some cases, criteria also work as indexes to measure the uncertainty in the analysis as they provided continuous index (e.g. from 0 to 1) rather than a discrete, yes or no, index . All of the analysis was done in the statistical software *R version 3.5.2 (2018-12-20)* with the packages *data.table*[@Packagedatatable:2019uh], *janitor*[@PackagejanitorSim:2018ur], *wesanderson*[@Packagewesanderson:2018vn], *rfishbase*[@PackagerfishbaseR:2019th], *R.matlab*[@PackageRmatlabRe:2018un], *sf*[@PackagesfSimpleF:2018vp], *sp*[@PackagespClasses:2019tn], *tidiverse*[@PackagetidyverseE:2017vq], *tidytext*[@PackagetidytextTe:2019vk], and *zoo*[@PackagezooSInfr:2019tm]. All code is available at https://github.com/jepa/FishForVisa.

@@

We use the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the 0.5 x 0.5 grid cells. We computed a *Species Index* by dividing the number of datasets that confirm presence of a species in a grid cell over the total amount of datasets (*n* = 3). The index is unit-less and represents a scale from 0.3 to 1 where 0.3 means only one dataset reports presence of the species in a grid cell, 0.6 represents two of three sources, and 1 means agreement between all datasets. For the analysis, we selected only those cases where *Species Index* = 1 to get a more conservative estimate of transboundary species.

@@

We estimated the economic contribution of transboundary species for each country using global ex-vessel price [@Tai2017]. The data includes ex-vessel price from multiple sources and a structured interpolation method (e.g. similar countries, species) to fill in data gaps [@Sumaila:2015uc]. The contribution of transboundary species was done by aggregating the catch of each species that happened within the EEZ of the countries sharing that species. For this study, we did not include catch nor revenue from areas beyond national jurisdiction. Finally, we use the catch data to determine the exploitation category of each species within a EEZ. Although this method has previously used to estimate stock status [@Grainger:1996tk], the categories presented here are intended to be seen as catch trends, and not status of each species. We only assessed species within each EEZ that had at least ten years of data between the first and last reported landings with at least five consecutive data years. Three categories were drawn from a method depending on the year's catch, the maximum catch over time and the minimum catch after the catch peak (**Table XX**) [@Kleisner:2011wn]. Finally, we reported the predominant category over the last ten years of catch data (2005-2014). We perform a series of one-way analysis of variance ( one-way ANOVA) to determine significant differences (statistically) between the means of different groups (e.g. categories, regions, species) regarding transboundary species and their contribution in terms of catch and revenue.

```{r Status_Table, eval = T, echo = F}

@@

Categories <- c("A - Growing Fishery",
                  "B - Stable Fishery",
                  "C - Fishery under 10% of Peak",
                  "No Status")

@@

data.frame(Categories,
           Rules) %>% 
  kable("latex", booktabs = T,
        caption = "Rules to determine the category of each transboundary species") %>%
  kable_styling(full_width = F) %>%
  column_spec(1,width = "3cm") %>% #wraps text in column2
  column_spec(2,width = "13cm") #%>% #wraps text in column2

@@

## Caveats

# Supplements

@@

It is expected that the number of gridcells where a species is present (*Area Index*) would directly influence whether or not a species is considered transboundary. Therefore, we tested the sensitivity of our results to the *Area Index* by estimating the number of transboundary species along a gradient of values (*S3 Index Area*). As expected, using a more relaxed value (e.g. the species is present within 10% or more of the EEZ) will result in more transboundary species giving a median of 40 EEZs shared per species, while a more strict value (e.g. 50%) will give a median of 4. For the current analysis we presented the results using a 25% *Area Index* threshold (*S3 Index Area* - darker histogram).

```{r Call_Sup_3, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Histogram of transboundary species using different values of Area Index. Darker histogram represents used value (25%) in the analysis"}

![**Histogram of transboundary species using different values of Area Index. Darker histogram represents used value (25%) in the analysis**](./Figures/S3_Area.png)

#References

<!-- # Recicled text (Not to be included) -->
